### 2024-09-24 paper/paper.qmd
Did some visualizations. Ready to write the paper!

---
title: "My title"
subtitle: "My subtitle if needed"
author: Cristina Su Lam
thanks: "Code and data supporting this analysis is available at: https://github.com/cristinaasu/TorontoTicketsIssued"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)       
library(kableExtra)      
library(here)            
library(scales)         
library(sf)             
library(ggmap)           
library(viridis)         
library(grid)            
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

```{r}
#| include: false
#| message: false
#| echo: false

# Read Cleaned Dataset
cleaned_tickets_data <- read_csv(
  file = here("data/analysis_data/cleaned_tickets_data.csv"),
  show_col_types = FALSE
)
```

Glimpse of Dataset

```{r}
#| label: tbl-ticketissuedata
#| tbl-cap: Sample of Toronto Tickets Issued Data
#| echo: false
#| warning: false
#| fig-align: center

# Ref from: 
# https://github.com/ThomasWilliamFox/toronto_child_care/blob/master/outputs/paper/toronto_child_care.qmd

head(cleaned_tickets_data, 6) |>
  kable(
    col.names = c("Ticket Count", " Offence Year", "Ticket Type", 
                  "Offence Category", "Age Group", "Division"),
    booktabs = TRUE
  )
```

Basic Summary Statistic Table for Ticket Count

```{r}
#| label: tbl-meansdtickets
#| tbl-cap: Mean and standard deviation of Ticket Count
#| echo: false
#| warning: false
#| fig-align: center

cleaned_tickets_data |>
  summarize(
    mean_ticket_count = sprintf("%.2f", mean(`Ticket Count`)),
    sd_ticket_count = sprintf("%.2f", sd(`Ticket Count`))
    ) |>
  kable(
    col.names = c("Mean Ticket Count", "SD of Ticket Count")
    )
```

Tickets Type vs Tickets Issued

```{r}
#| label: tbl-ticketcatg
#| tbl-cap: Frequency of Ticket Categories
#| echo: false
#| warning: false
#| fig-align: center
# Bar chart of ticket categories and their frequencies
cleaned_tickets_data |> 
  ggplot(aes(x = `Ticket Type`)) +
  geom_bar(fill = "steelblue") + 
  theme_minimal() +
  labs(x = "Offence Category",
       y = "Count of Tickets") +
  theme(axis.text.x = element_text(hjust = 1))
```

Offence Category vs Tickets Issued

```{r}
#| label: tbl-offencebyticket
#| tbl-cap: Stacked Bar Chart of Offence Categories by Ticket Type
#| echo: false
#| warning: false
#| fig-align: center
# Stacked bar chart of Offence Category and Tickets Issued
cleaned_tickets_data |> 
  ggplot(aes(x = `Offence Category`, fill = `Ticket Type`)) + 
  geom_bar(position = "stack") + 
  theme_minimal() + 
  labs(x = "Count of Tickets", 
       y = "Offence Category", 
       fill = "Ticket Type") + 
  coord_flip() +  # Flip the axes for a horizontal layout
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  
```

Ticket Counts by Ticket Type (Doughnut Chart)

```{r}
#| label: tbl-ticketcount
#| tbl-cap: Distribution of Tickets by Type
#| echo: false
#| warning: false
#| fig-align: center
# Summarize the ticket counts by ticket type
ticket_type_summary <- cleaned_tickets_data |> 
  group_by(`Ticket Type`) |> 
  summarise(Total_Tickets = sum(`Ticket Count`)) |> 
  ungroup() |> 
  mutate(Percentage = Total_Tickets / sum(Total_Tickets) * 100)

# Create the doughnut chart
ggplot(ticket_type_summary, aes(x = 2, y = Total_Tickets, fill = `Ticket Type`)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +  
  theme_void() +  
  geom_text(aes(label = paste0(Total_Tickets, " (", round(Percentage, 1), "%)")),
            position = position_stack(vjust = 0.5), color = "black") + 
  scale_fill_brewer(palette = "Set3") + 
  xlim(0.5, 2.5)  
```

Age Group vs Tickets Issued
```{r}
#| label: tbl-age
#| tbl-cap: Stacked Bar Chart of Offence Categories by Age Group
#| echo: false
#| warning: false
#| fig-align: center
cleaned_tickets_data |> 
  ggplot(aes(x = `Offence Category`, fill = `Age Group`)) + 
  geom_bar(position = "stack") + 
  theme_minimal() + 
  labs(x = "Count of Tickets", 
       y = "Offence Category", 
       fill = "Age Group") + 
  coord_flip() +  # Flip the axes for a horizontal layout
  theme(axis.text.y = element_text(angle = 0, hjust = 1)) 
```

Annual Patterns
```{r}
#| label: tbl-yearlyp
#| tbl-cap: Ticket Counts by Offence Category and Year
#| echo: false
#| warning: false
#| fig-align: center
# Summarizing data to get ticket counts by offence category and year
ticket_counts <- cleaned_tickets_data %>%
  group_by(`Offence Category`, `Offence Year`) %>%
  summarize(`Ticket Count` = sum(`Ticket Count`), .groups = 'drop')

# Creating the line graph
ticket_counts |> 
  ggplot(aes(x = `Offence Category`, y = `Ticket Count`,
             color = as.factor(`Offence Year`), 
             group = `Offence Year`, linetype = as.factor(`Offence Year`))) + 
  geom_line(size = 1.2) +  # Increase line width
  geom_point(size = 3) +   # Optional: Adjust point size
  theme_minimal() + 
  labs(x = "Offence Category", 
       y = "Ticket Count", 
       color = "Year", 
       linetype = "Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
```

Geographic Map divided by divisions and shows #tickets issued per division

```{r}
#| label: tbl-map
#| tbl-cap: Total Tickets Issued by Division
#| echo: false
#| warning: false
#| fig-align: center
# 1. Summarize ticket data by division
tickets_summary <- cleaned_tickets_data %>%
  group_by(Division) %>%
  summarise(total_tickets = sum(`Ticket Count`, na.rm = TRUE))

# 2. Merge the summarized tickets data with the GeoJSON division data
division_with_tickets <- division_data %>%
  left_join(tickets_summary, by = c("DIV" = "Division"))

# 3. Plot the map with custom colors
ggplot(data = division_with_tickets) +
  geom_sf(aes(fill = total_tickets), color = "black") +  
  geom_sf_text(aes(label = paste0(DIV, "\n", total_tickets)),  
               size = 2, color = "black") + 
  scale_fill_gradientn(colors = c("#FFCCCB", "#FFDDC1", "#FFE4B5", "#D3FFCE", "#B2E0E0"),
                       labels = scales::comma) + 
  labs(fill = "Total Tickets") +  
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5), 
    axis.title = element_blank(),             
    axis.text.x = element_blank(),            
    axis.text.y = element_blank(),            
    axis.ticks = element_blank(),              
    panel.grid.major = element_blank(),       
    panel.grid.minor = element_blank()         
  ) +
  # Add rectangle around the entire plot (map, title, and legend)
  annotation_custom(
    grob = rectGrob(gp = gpar(fill = NA, col = "black", lwd = 1)),
    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  )
```

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

# References