### 2024-09-24 paper/paper.qmd
First Draft

---
title: "Offenses and Enforcement: A Study of Toronto’s Ticket Issuance Patterns"
author: Cristina Su Lam
thanks: "Code and data supporting this analysis is available at: https://github.com/cristinaasu/TorontoTicketsIssued"
date: today
date-format: long
abstract: "This paper analyzes Toronto’s ticket issuance data from 2018 to 2023 to uncover patterns in traffic enforcement across various offense categories and demographic groups. The analysis reveals that certain offense types, such as [X] and [Y], consistently result in higher ticket counts, with significant geographic concentrations in specific city divisions. Trends in ticket issuance by age group and year also show fluctuations, with noticeable increases in enforcement activity during certain periods. These findings offer valuable insights for guiding targeted law enforcement policies and improving the allocation of resources across the city."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)       
library(kableExtra)      
library(here)            
library(scales)         
library(sf)             
library(ggmap)           
library(viridis)         
library(grid)  
library(gt)

```

# Introduction

Traffic violations play a significant role in shaping urban transportation systems, impacting both road safety and public behavior. In cities like Toronto, where traffic congestion and mobility issues are growing concerns, effective enforcement of traffic laws is essential to maintain order and safety. The issuance of tickets for violations such as speeding, distracted driving, and aggressive driving is a critical tool for law enforcement to manage these challenges. Understanding how these tickets are distributed across geographic areas and demographic groups can help policymakers ensure that enforcement is both fair and effective.

Recent studies suggest that specific offense types, such as speeding and distracted driving, contribute disproportionately to road safety hazards. While these violations are well-recognized, there is limited research on how the enforcement of these offenses is distributed across different city divisions or demographic groups. Ticket issuance data can provide insight into these trends, revealing whether certain areas or age groups are more likely to be ticketed. This information is vital for creating targeted strategies to address high-risk behaviors. Between 2011 and 2020, road safety in Canada improved in terms of overall rates of fatalities and serious injuries, but speeding (25%) and distracted driving (21%) remained leading contributors to fatal accidents in 2020【Cite: https://tc.canada.ca/en/road-transportation/publications/road-safety-canada-2020 】.

This paper focuses on ticket issuance data in Toronto from 2018 to 2023 to examine enforcement trends before and after the COVID-19 pandemic, which significantly impacted traffic volume, public behavior, and law enforcement activities. The analysis will examine offense categories, geographic distribution, and demographic patterns, uncovering key trends in ticket issuance across the city. This study identifies which types of offenses consistently result in higher ticket counts, examines their concentration in specific city divisions, and explores how age demographics influence enforcement. It also tracks yearly fluctuations to assess broader trends in enforcement over time【Cite: Studies on pandemic-related changes in urban traffic enforcement】.

These findings provide valuable insights for guiding law enforcement strategies and resource allocation across Toronto. By identifying areas and demographic groups with higher rates of traffic violations, policymakers can implement more targeted interventions to improve traffic safety and ensure equitable enforcement practices. The paper is structured as follows: Section Two discusses the data and methodology, Section Three presents the findings, and Section Four explores the implications and suggests areas for further research.

# Data Analysis {#sec-data}

```{r}
#| include: false
#| message: false
#| echo: false

# Read Cleaned Dataset
cleaned_tickets_data <- read_csv(
  file = here("data/analysis_data/cleaned_tickets_data.csv"),
  show_col_types = FALSE
)

# Read GeoJSON data correctly
cleaned_division_data <- st_read(here("data/analysis_data/cleaned_division_data.geojson"))
```

To explore patterns of ticket issuance in Toronto from 2018 to 2023, comprehensive data were acquired from the Toronto Open Data catalogue using the OpenDataToronto R package (Gelfand 2022). This dataset is continually updated, with the latest records up to August 2nd, 2024, ensuring relevance and accuracy for this analysis.

The initial dataset comprised various entries which included data points categorized by division, offense year, ticket type, and demographic details. To refine and optimize the dataset for this analysis, entries with unspecified division locations marked as "NSA" (No Specified Address), which generally denote occurrences outside the jurisdictional bounds of Toronto or those without a confirmed location, were excluded. Similarly, entries under the "Unknown" category within the age group variable were also omitted to maintain the integrity and reliability of the demographic analysis. Post-cleanup, the dataset was reduced to 15,672 observations, all meticulously documented from 2018 through 2023.

The statistical analysis was conducted using the R programming language, utilizing its powerful tidyverse suite (Wickham et al., 2019) for data manipulation and ggplot2 for visualization (Wickham, 2016). The key focus was to ascertain the distribution and variability of ticket issuance across different years and analyze demographic trends.

The cleaned dataset revealed a total of 1,145,119 tickets issued over the six-year period, averaging 190,853.2 tickets per year. The standard deviation of 14,082.49 highlighted a notable variation in the number of tickets issued annually, suggesting influences of external factors or changes in enforcement intensity. This variability warranted a deeper exploration to understand annual trends and their implications on urban management and policy enforcement strategies.

```{r}
#| label: tbl-ticketissuedata
#| tbl-cap: Sample of Toronto Tickets Issued Data
#| echo: false
#| warning: false
#| fig-align: center

# Ref from: 
# https://github.com/ThomasWilliamFox/toronto_child_care/blob/master/outputs/paper/toronto_child_care.qmd

cleaned_tickets_data |>  
  head(6) |> 
  kable( 
    col.names = c("Offence Year", "Division", "Ticket Type", 
                  "Offence Category", "Age Group", "Ticket Count"), 
    booktabs = TRUE 
  )
```

```{r}
#| label: tbl-meansdtickets
#| tbl-cap: Summary Statistics for Ticket Issuance in Toronto from 2018-2023
#| echo: false
#| warning: false
#| fig-align: center

# Calculate total tickets issued
total_tickets <- cleaned_tickets_data %>%
  summarize(total = sum(`Ticket Count`)) %>%
  pull(total)

# Calculate average tickets issued per year
average_tickets_per_year <- cleaned_tickets_data %>%
  group_by(year = `Offence Year`) %>% 
  summarize(total = sum(`Ticket Count`)) %>%
  summarize(average = mean(total))

# Calculate standard deviation of yearly tickets issued
std_dev_yearly_tickets <- cleaned_tickets_data %>%
  group_by(year = `Offence Year`) %>%
  summarize(total = sum(`Ticket Count`)) %>%
  summarize(sd = sd(total))

# Create summary table with formatted values
summary_table <- data.frame(
  `Summary Statistic` = c("Total Tickets Issued", 
                "Mean Tickets Issued per Year", 
                "Standard Deviation of Yearly Tickets Issued"),
  Value = c(format(total_tickets, big.mark = ",", scientific = FALSE), 
            format(round(average_tickets_per_year$average, 2), big.mark = ",", scientific = FALSE), 
            format(round(std_dev_yearly_tickets$sd, 2), big.mark = ",", scientific = FALSE))
)

# Print the summary table using kable
kable(summary_table, format = "pipe")

```

## Offence Categories {#subsec1-data}

Analysis of ticket categories reveals marked variations in enforcement across different offences. "Speeding" and "Other HTA" offences dominate ticket issuance, highlighting rigorous enforcement likely aimed at reducing high-speed incidents and enhancing road safety. Notably, the data shows that "Speeding" consistently incurs the highest number of tickets, underscoring its priority in traffic law enforcement agendas.

Conversely, tickets for "Distracted Driving" and "Aggressive Driving" are less frequent, which might suggest areas for increased enforcement. Given the dangerous nature of these offenses, particularly "Distracted Driving," which is increasingly problematic with the ubiquity of mobile devices, there may be an opportunity to intensify focus here. This could include more stringent monitoring techniques or public awareness campaigns targeting these specific behaviors.

Tickets Type vs Tickets Issued
```{r}
#| label: tbl-ticketcatg
#| tbl-cap: Frequency of Ticket Categories
#| echo: false
#| warning: false
#| fig-align: center
# Bar chart of ticket categories and their frequencies
cleaned_tickets_data |>  
  group_by(`Ticket Type`) |>          # Group by Ticket Type
  summarize(total_tickets = sum(`Ticket Count`, na.rm = TRUE)) |>  # Sum the Ticket Count
  ggplot(aes(x = `Ticket Type`, y = total_tickets)) +  # Use summed values for y-axis
  geom_bar(stat = "identity", fill = "steelblue") +  # Use stat = "identity" to plot summed values
  scale_y_continuous(labels = label_comma()) +  # Format y-axis labels with commas
  theme_minimal() + 
  labs(x = "Ticket Type",
       y = "Total Count of Tickets") + 
  theme(axis.text.x = element_text(hjust = 1))
```

Offence Category vs Tickets Issued

```{r}
#| label: tbl-offencebyticket
#| tbl-cap: Stacked Bar Chart of Offence Categories by Ticket Type
#| echo: false
#| warning: false
#| fig-align: center
# Stacked bar chart of Offence Category and Tickets Issued
cleaned_tickets_data |>  
  group_by(`Offence Category`, `Ticket Type`) |>  # Group by Offence Category and Ticket Type
  summarize(total_tickets = sum(`Ticket Count`, na.rm = TRUE)) |>  # Sum the Ticket Count
  ggplot(aes(x = `Offence Category`, y = total_tickets, fill = `Ticket Type`)) + 
  geom_bar(stat = "identity", position = "stack") +  # Use stat = "identity" for totals
  theme_minimal() + 
  labs(x = "Count of Tickets", 
       y = "Offence Category", 
       fill = "Ticket Type") + 
  coord_flip() +  # Flip the axes for a horizontal layout
  theme(axis.text.y = element_text(angle = 0, hjust = 1)) +
  scale_y_continuous(labels = scales::label_comma())  # Format y-axis labels with commas
```

Ticket Counts by Ticket Type (Doughnut Chart)
```{r}
#| label: tbl-ticketcount
#| tbl-cap: Distribution of Tickets by Type
#| echo: false
#| warning: false
#| fig-align: center
# Summarize the ticket counts by ticket type
ticket_type_summary <- cleaned_tickets_data |> 
  group_by(`Ticket Type`) |> 
  summarise(Total_Tickets = sum(`Ticket Count`)) |> 
  ungroup() |> 
  mutate(Percentage = Total_Tickets / sum(Total_Tickets) * 100)

# Create the doughnut chart
ggplot(ticket_type_summary, aes(x = 2, y = Total_Tickets, fill = `Ticket Type`)) + 
  geom_bar(stat = "identity", width = 1, color = "white") + 
  coord_polar("y", start = 0) + 
  theme_void() + 
  geom_text(aes(label = paste0(scales::label_comma()(Total_Tickets), " (", round(Percentage, 1), "%)")),
            position = position_stack(vjust = 0.5), color = "black") + 
  scale_fill_brewer(palette = "Set3") + 
  xlim(0.5, 2.5) 
```
## Age Group {#subsec2-data}
The data indicates a significant disparity in ticket issuance between adults and youths, with adults receiving the majority of tickets. This distribution likely mirrors the demographic composition of active drivers but also raises questions about the driving behavior between different age groups.

An in-depth analysis into the types of offenses by age group could unveil targeted opportunities for intervention. Educational programs, especially for new drivers, could be crucial in instilling good driving habits early. Additionally, exploring the impact of graduated licensing could yield insights into its effectiveness in mitigating youth-related traffic violations.

```{r}
#| label: tbl-age
#| tbl-cap: Stacked Bar Chart of Offence Categories by Age Group
#| echo: false
#| warning: false
#| fig-align: center
# Stacked bar chart of Offence Category and Tickets Issued by Age Group
cleaned_tickets_data |> 
  group_by(`Offence Category`, `Age Group`) |>  # Group by Offence Category and Age Group
  summarize(total_tickets = sum(`Ticket Count`, na.rm = TRUE)) |>  # Sum the Ticket Count
  ggplot(aes(x = `Offence Category`, y = total_tickets, fill = `Age Group`)) + 
  geom_bar(stat = "identity", position = "stack") +  # Use stat = "identity" for totals
  theme_minimal() + 
  labs(x = "Count of Tickets", 
       y = "Offence Category", 
       fill = "Age Group") + 
  coord_flip() +  # Flip the axes for a horizontal layout
  theme(axis.text.y = element_text(angle = 0, hjust = 1)) +
  scale_y_continuous(labels = scales::label_comma())  # Format y-axis labels with commas
```
## 2.3 Yearly Patterns {#subsec3-data}
Reviewing ticket issuance over several years shows fluctuations, with a noticeable peak in 2023. Such variations could reflect changes in enforcement intensity, alterations in traffic laws, or broader socio-economic trends affecting driving patterns.

Examining the influence of policy changes, economic factors, or even public events could explain some of the annual variations observed. This could be particularly relevant for understanding the impact of global events like the COVID-19 pandemic on traffic patterns and enforcement.

```{r}
#| label: tbl-yearlyp
#| tbl-cap: Ticket Counts by Offence Category and Year
#| echo: false
#| warning: false
#| fig-align: center
ticket_counts <- cleaned_tickets_data %>%
  group_by(`Offence Category`, `Offence Year`) %>%
  summarize(`Ticket Count` = sum(`Ticket Count`), .groups = 'drop')

# Convert your data into a wider format
ticket_counts_wide <- ticket_counts %>%
  pivot_wider(names_from = `Offence Year`, values_from = `Ticket Count`, values_fill = list(`Ticket Count` = 0))

# Calculate totals for each year
totals <- ticket_counts_wide %>%
  summarize(across(-`Offence Category`, sum, na.rm = TRUE)) %>%
  mutate(`Offence Category` = "Total")

# Append the totals row
ticket_counts_wide <- bind_rows(ticket_counts_wide, totals)

# Create the table using gt
ticket_table <- gt(ticket_counts_wide) %>%
  cols_label(
    `2018` = "2018",
    `2019` = "2019",
    `2020` = "2020",
    `2021` = "2021",
    `2022` = "2022",
    `2023` = "2023"
  ) %>%
  tab_header(
    title = "Tickets",
    subtitle = "Tickets Issued by Category and Year"
  ) %>%
  fmt_number(
    columns = vars(`2018`, `2019`, `2020`, `2021`, `2022`, `2023`),
    decimals = 0
  ) %>%
  tab_options(
    table.width = pct(100),
    column_labels.font.size = "small"
  )

# View the table
print(ticket_table)

```

## Geographic Distribution {#subsec4-data}

Ticket distribution by geographic division highlights areas with varied enforcement levels. The highest concentrations of tickets in divisions D41, D55, and D14 suggest that these areas could be high traffic zones or regions with higher incidences of traffic violations. This distribution may reflect strategic placement of traffic law enforcement resources in areas perceived as high-risk.

Further analysis could correlate high ticket areas with external factors such as traffic volume, accident rates, or demographic data. Understanding these correlations can help in optimizing resource allocation, ensuring that enforcement efforts are both effective and efficient.

```{r}
#| label: tbl-map
#| tbl-cap: Total Tickets Issued by Division
#| echo: false
#| warning: false
#| message: false
#| fig-align: center
# 1. Summarize ticket data by division 
tickets_summary <- cleaned_tickets_data %>% 
  group_by(Division) %>% 
  summarise(total_tickets = sum(`Ticket Count`, na.rm = TRUE))

# 2. Merge the summarized tickets data with the GeoJSON division data 
division_with_tickets <- cleaned_division_data %>% 
  left_join(tickets_summary, by = c("Division" = "Division"))

# 3. Plot the map with custom colors 
ggplot(data = division_with_tickets) + 
  geom_sf(aes(fill = total_tickets), color = "black") +   
  geom_sf_text(aes(label = paste0(Division, "\n", scales::comma(total_tickets))),   # Add commas here
               size = 2, color = "black") +  
  scale_fill_gradientn(colors = c("#FFCCCB", "#FFDDC1", "#FFE4B5", "#D3FFCE", "#B2E0E0"),
                       labels = scales::comma) +  
  labs(fill = "Total Tickets") +   
  theme_minimal() +  
  theme( 
    plot.title = element_text(hjust = 0.5),  
    axis.title = element_blank(),             
    axis.text.x = element_blank(),             
    axis.text.y = element_blank(),             
    axis.ticks = element_blank(),               
    panel.grid.major = element_blank(),        
    panel.grid.minor = element_blank()          
  ) + 
  # Add rectangle around the entire plot (map, title, and legend) 
  annotation_custom( 
    grob = rectGrob(gp = gpar(fill = NA, col = "black", lwd = 1)), 
    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf 
  )
```

# Discussion {#sec-discussion}
## Results and Implications {#subsec1-discussion}
This study highlights the complexity and challenges of traffic law enforcement in a major urban center. The predominance of speeding and other HTA offenses suggests that while enforcement measures are effective in capturing these violations, ongoing issues persist that require continuous effort and adaptation of strategies.

Implications: Enhancing enforcement for less frequently ticketed but highly risky behaviors such as distracted and aggressive driving could further improve road safety. Also, differential enforcement strategies considering the unique characteristics of each division may lead to more efficient use of resources and greater compliance with traffic laws.

## Further Areas of Exploration {#subsec2-discussion}
The initial findings suggest several avenues for deeper investigation. Analyzing the causes for the geographic and demographic patterns in ticket issuance could provide insights into underlying factors influencing these trends.

Suggestions: Integrating data on traffic flow and vehicle density with ticket issuance could offer a more comprehensive understanding of the effectiveness of current traffic law enforcement strategies. Additionally, employing advanced analytical techniques, such as predictive analytics or machine learning, could forecast trends and inform proactive traffic management strategies.

This paper offers a foundational exploration into the patterns of ticket issuance within Toronto, providing a base from which policymakers and researchers can further investigate and improve traffic law enforcement strategies. By continuing to analyze these patterns and their broader implications, strategic improvements can be formulated to enhance both compliance and overall road safety across the city.

This expanded discussion provides a thorough analysis of your data, ensuring each section of the paper is robustly supported by the visualizations and the underlying data you've provided.

\newpage

\appendix

# Appendix {.unnumbered}

# References